---
title: "Survival Analysis"
author: "Hazlienor"
date: "2023-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Survival Analysis
## Workflow
1. prepare environment
2. read data
3. data exploration
4. Kaplan-Meier Analysis

## prepare environment
```{r}
library(tidyverse)
library(psych)
library(corrplot)
library(knitr)
library(broom)
library(tidyr)
library(survival)
library(survminer)
library(dplyr)
library(lubridate)
library(haven)
library(here)
library(gtsummary)
```

## read data
```{r}
stroke <- read_dta(here('stroke_outcome.dta'))
```

## data exploration
```{r}
view(stroke)
str(stroke)
```
```{r}
stroke <- stroke %>% mutate_if(is.labelled, ~as_factor(.))
```

Exploratory data analysis
```{r}
summary(stroke)
```
Format date
```{r}
stroke <- stroke %>% mutate(dur = stroke$doa %--% stroke$dod) %>%
  mutate(dur = as.duration(dur))

# convert to days
stroke <- stroke %>% mutate(dur_days = dur /ddays(1))
```

summarize data overall
```{r}
stroke %>% 
  select(-id, -event_s, -race2, -dur, -dur_days) %>%
  tbl_summary()  %>%
  as_gt()
```



describe data by event (numerical)
```{r}
stroke %>% group_by(event) %>% summarise(mean.age = mean(age), sd.age = sd(age),
                                           mean.gcs = mean(gcs), sd.gcs = sd(gcs),
                                           mean.sbp = mean(sbp), sd.sbp = sd(sbp),
                                           mean.dbp = mean(dbp), sd.dbp = sd(dbp))
```

describe data by event (categorical)
```{r}
stroke %>% count(event, dm2)  
```

```{r}
stroke %>% count(event, hpt2)  
```

```{r}
stroke %>% count(event, race2)  
```

## Kaplan Meier

Kaplan Meier - to estimate survival probability

conduct KM analysis
```{r}
KM1 <- survfit(Surv(time = dur_days, event == 'dead') ~ 1, 
               type = "kaplan-meier", data = stroke)
summary(KM1)
```
```{r}
survfit(Surv(time = dur_days, event == 'dead') ~ 1, 
               type = "kaplan-meier", data = stroke)

# the estimated median survival time is 22 days
```

probability of surviving at day 2
```{r}
(1-(4/224))*0.935
## calculation at 2 days = 1-(4/224)*0.935 (previous day probability)
```


plot KM
```{r}
ggsurvplot(KM1, data = stroke, risk.table = TRUE, linetype = c(1,2), pval = TRUE)
```
Median line
```{r}
  ggsurvplot(
  KM1, 
  conf.int = FALSE, 
  surv.median.line = "hv",
  xlab = "Days", 
  ylab = "Survival probability", 
  break.time.by = 20,
  risk.table = T)
```


KM based on gender
```{r}
KM1.g <- survfit(Surv(time = dur_days, event == 'dead') ~ sex, 
                     type = "kaplan-meier", data = stroke)
summary(KM1.g)
```
plot KM by gender
```{r}
ggsurvplot(KM1.g, data = stroke, risk.table = TRUE, 
           linetype = c(1,2), pval = TRUE)
```
median survival days based on DM
```{r}
KM1.dm
```


KM by DM
```{r}
KM1.dm <- survfit(Surv(time = dur_days, event == 'dead') ~ dm2, 
                     type = "kaplan-meier", data = stroke)
summary(KM1.dm)
```
```{r}
ggsurvplot(KM1.dm, data = stroke, risk.table = TRUE, 
           linetype = c(1,2), pval = TRUE)
```
KM by HPT
```{r}
KM1.hpt <- survfit(Surv(time = dur_days, event == 'dead') ~ hpt2, 
                     type = "kaplan-meier", data = stroke)
summary(KM1.hpt)
```

```{r}
ggsurvplot(KM1.hpt, data = stroke, risk.table = TRUE, 
           linetype = c(1,2), pval = TRUE)
```
Estimate Survival Function - overall
```{r}
quantile(KM1, probs = c(0.25, 0.50, 0.75))
```

Estimate Survival Function - by DM
```{r}
quantile(KM1.dm, probs = c(0.25, 0.50, 0.75))
```
Estimate Survival Probability
```{r}
stroke %>% group_by(event) %>% 
  summarize(min.dur = min(dur_days), max.dur = max(dur_days))
```
Estimate Survival Probability at different time
```{r}
summary(KM1, times = c(35, 40, 45))
```
Estimate Survival Probability at different time based on DM status
```{r}
summary(KM1.dm, times = c(15, 30, 40))
```

## Comparing survival rate between groups

survival rate between gender - log rank
```{r}
logrank.sex <- survdiff(Surv(time = dur_days, event == 'dead') ~ sex, 
                        data = stroke, rho = 0)
logrank.sex
```


survival rate between gender - peto-peto
```{r}
peto.sex <- survdiff(Surv(time = dur_days, event == 'dead') ~ sex, 
                     data = stroke, rho = 1)
peto.sex
```
survival rate between DM/non-DM - log rank
```{r}
logrank.dm <- survdiff(Surv(time = dur_days, event == 'dead') ~ dm2, 
                        data = stroke, rho = 0)
logrank.dm
```

survival rate between DM/non-DM - peto-prentice
```{r}
peto.dm <- survdiff(Surv(time = dur_days, event == 'dead') ~ dm2, 
                     data = stroke, rho = 1)
peto.dm

# there was a significant difference in overall survival probability according to DM status among stroke patient, with a p-value of p <0.001
```

## Cox Proportionate Hazard

